---
title: 'Case Study 1'
author: 'Austin Webb'
date: '2022-06-12'
output: html_document
---

# Read in the packages needed
```{r}
library(dplyr) # yes
library(tidyr) # yes
library(caret) # yes
library(ggplot2) # yes

# library(plotly)
# library(stringi)
# library(rvest) # html_table, html_node
# library(RCurl) # getURL
# library(tm) # text mining library provides the stopwords() function
# library(plyr)
# library(jsonlite)
# library(tidyverse)
# library(class)
# library(e1071)
# library(fastDummies)
# library(cluster)
```

# Data import
```{r}
# Read in the beer data
beer <- read.csv(file.choose(), header = TRUE)
head(beer)
summary(beer)

# Read in the brewery data
brewery <- read.csv(file.choose(), header = TRUE)
head(brewery)
summary(brewery)
```


# Question 1
## How many breweries are present in each state?
```{r}
# This visually looks cleaner
table(brewery$State)
state_count <- data.frame(table(brewery$State))

# Same info different style
brewery %>% count(State)



```


# Question 2
## Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
```{r}
# Merge the two dataframes
beer_brewery_merged <- merge(beer, brewery, by.x='Brewery_id', by.y = 'Brew_ID')

# Print the first six observations
head(beer_brewery_merged, 6)

# Print the last six observations
tail(beer_brewery_merged, 6)

```


# Question 3
## Address the missing values in each column.
```{r}
# Overview of what we are looking at
summary(beer_brewery_merged)
# we see that we have missing values in ABV (62) and IBU (1005)

# Check to see if this adds up
sum(is.na(beer_brewery_merged))

# Nice check to see missing values in each column
sapply(beer_brewery_merged, function(x) sum(is.na(x)))

# Make copy of dataframe
beer_brewery_merged_mean <- beer_brewery_merged

# Replace nan in ABV column with mean
beer_brewery_merged_mean$ABV[is.na(beer_brewery_merged_mean$ABV)] <- mean(beer_brewery_merged_mean$ABV, na.rm = TRUE)
# Replace nan in IBU column with mean
beer_brewery_merged_mean$IBU[is.na(beer_brewery_merged_mean$IBU)] <- mean(beer_brewery_merged_mean$IBU, na.rm = TRUE)

# Check to see if the nans are gone
summary(beer_brewery_merged_mean)

# Double check to see if we missed anything
sum(is.na(beer_brewery_merged_mean))

# Nice check to see missing values in each column
sapply(beer_brewery_merged_mean, function(x) sum(is.na(x)))



# Drop the observations where there is a nan in ABV or IBU
beer_brewery_merged_dropped_nan <- filter(beer_brewery_merged, !is.na(ABV))
beer_brewery_merged_dropped_nan <- filter(beer_brewery_merged_dropped_nan, !is.na(IBU))


# Check to see if the nans are gone
summary(beer_brewery_merged_dropped_nan)

# Double check to see if we missed anything
sum(is.na(beer_brewery_merged_dropped_nan))

# Nice check to see missing values in each column
sapply(beer_brewery_merged_dropped_nan, function(x) sum(is.na(x)))


```


# Question 4
## Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
### Use the mean for nan dataset
```{r}
# Find mean, min, max, median, and standard deviation ABV for each state 
beer_brewery_merged_mean %>%
  group_by(State) %>% 
  summarise(Mean = mean(ABV), 
            Min = min(ABV), 
            Max = max(ABV),
            Median = median(ABV), 
            Std = sd(ABV))

# Find mean, min, max, median, and standard deviation IBU for each state 
beer_brewery_merged_mean %>%
  group_by(State) %>% 
  summarise(Mean = mean(IBU), 
            Min = min(IBU), 
            Max = max(IBU),
            Median = median(IBU), 
            Std = sd(IBU))

# Just find the median of ABV for each state
ABV_Medians_mean <- beer_brewery_merged_mean %>%
  group_by(State) %>% 
  summarise(Median = median(ABV))

# Just find the median of IBU for each state
IBU_Medians_mean <- beer_brewery_merged_mean %>%
  group_by(State) %>% 
  summarise(Median = median(IBU))



```

```{r}
# Plot the ABV medians by state
abv_plot_mean <- ggplot(ABV_Medians, 
                   aes(x = State, 
                       y = Median, 
                       fill = State)) +
  geom_bar(width = 0.4, 
           position = position_dodge(width = 0.05), 
           stat = 'identity') +
  #coord_flip()  +
  ggtitle('Median ABV by State') +
  geom_hline(yintercept = mean(ABV_Medians$Median))

# Add labels
abv_plot_mean + labs(title = 'Median ABV by State', 
                x = 'State', 
                y = 'Median')
abv_plot_mean + theme(axis.text = element_text(size = 8))

abv_plot_mean + theme(axis.text.x = element_text(angle = 90, 
                                            hjust = .25, 
                                            vjust = .25,
                                            size = 6))
```





```{r}

# plot the IBU medians by state
ibu_plot <- ggplot(IBU_Medians, 
                   aes(x = State, 
                       y = Median, 
                       fill = State)) +
  geom_bar(width = 0.4, 
           position = position_dodge(width = 0.05),
           stat = 'identity') +
 # coord_flip() +
  ggtitle('Median IBU by State') +
  geom_hline(yintercept = mean(IBU_Medians$Median))

# Add labels
ibu_plot + labs(title = 'Median IBU by State', 
                x = 'State', 
                y = 'Median')
ibu_plot + theme(axis.text = element_text(size = 8)) 

ibu_plot + theme(axis.text.x = element_text(angle = 90, 
                                            hjust = .25, 
                                            vjust = .25,
                                            size = 6))


```



```{r}

# Merge the medians to create one df with state medians for ABV and IBU 

ABV_IBU_merged <- merge(ABV_Medians, 
                        IBU_Medians, 
                        by = 'State')
# Rename the columns
ABV_IBU_merged <- rename(ABV_IBU_merged, 
                         'ABV Medians' = 'Median.x')
ABV_IBU_merged <- rename(ABV_IBU_merged, 
                         'IBU Medians' = 'Median.y')

# Pivot the data to make it easier to plot
to_plot_2 <- pivot_longer(ABV_IBU_merged, 
                          cols = c('ABV Medians', 
                                   'IBU Medians'), 
                          names_to = 'Acronym', 
                          values_to = 'Medians')

# Plot the data
combined_plot <- ggplot(to_plot_2, 
                        aes(x = State, 
                            y = Medians, 
                            fill = Acronym)) +
  geom_bar(position='dodge', 
           stat='identity')
# Change the orientation of the text on the x axis
combined_plot + theme(axis.text.x = element_text(angle = 90, 
                                                 vjust = 0.5, 
                                                 hjust = 1))
```

### Use the dropped for nan dataset
```{r}
# Find mean, min, max, median, and standard deviation ABV for each state 
beer_brewery_merged_dropped_nan %>%
  group_by(State) %>% 
  summarise(Mean = mean(ABV), 
            Min = min(ABV), 
            Max = max(ABV),
            Median = median(ABV), 
            Std = sd(ABV))

# Find mean, min, max, median, and standard deviation IBU for each state 
beer_brewery_merged_dropped_nan %>%
  group_by(State) %>% 
  summarise(Mean = mean(IBU), 
            Min = min(IBU), 
            Max = max(IBU),
            Median = median(IBU), 
            Std = sd(IBU))

# Just find the median of ABV for each state
ABV_Medians_dropped <- beer_brewery_merged_dropped_nan %>%
  group_by(State) %>% 
  summarise(Median = median(ABV))

# Just find the median of IBU for each state
IBU_Medians_dropped <- beer_brewery_merged_dropped_nan %>%
  group_by(State) %>% 
  summarise(Median = median(IBU))



```

#### Maybe sort the data??? -> when coparing the two plots its easier to see if the axes are the same
```{r}
# Plot the ABV medians by state
abv_plot_dropped <- ggplot(ABV_Medians_dropped, 
                   aes(x = State, 
                       y = Median, 
                       fill = State)) +
  geom_bar(width = 0.4, 
           position = position_dodge(width = 0.05), 
           stat = 'identity') +
  #coord_flip()  +
  ggtitle('Median ABV by State') +
  geom_hline(yintercept = mean(ABV_Medians_dropped$Median))

# Add labels
abv_plot_dropped + labs(title = 'Median ABV by State', 
                x = 'State', 
                y = 'Median')
abv_plot_dropped + theme(axis.text = element_text(size = 8))

abv_plot_dropped + theme(axis.text.x = element_text(angle = 90, 
                                            hjust = .25, 
                                            vjust = .25,
                                            size = 6))
```


```{r}

# plot the IBU medians by state
ibu_plot_dropped <- ggplot(IBU_Medians_dropped, 
                   aes(x = State, 
                       y = Median, 
                       fill = State)) +
  geom_bar(width = 0.4, 
           position = position_dodge(width = 0.05),
           stat = 'identity') +
 # coord_flip() +
  ggtitle('Median IBU by State') +
  geom_hline(yintercept = mean(IBU_Medians$Median))

# Add labels
ibu_plot_dropped + labs(title = 'Median IBU by State', 
                x = 'State', 
                y = 'Median')
ibu_plot_dropped + theme(axis.text = element_text(size = 8)) 

ibu_plot_dropped + theme(axis.text.x = element_text(angle = 90, 
                                            hjust = .25, 
                                            vjust = .25,
                                            size = 6))


```









# Question 5
## Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
### Use the mean for nan dataset
```{r}

# ABV
max(beer_brewery_merged_mean$ABV)
# This shows the max ABV is 0.128
max_abv_location_mean <- which.max(beer_brewery_merged_mean$ABV)
# This shows that the max ABV is found in index 375
beer_brewery_merged_mean %>% slice(max_abv_location_mean)
# This shows that the max ABV is in Boulder, Colorado

# IBU
max(beer_brewery_merged_mean$IBU)
# This shows the max ABV is 138
max_ibu_location_mean <- which.max(beer_brewery_merged_mean$IBU)
# This shows that the max ABV is found in index 1857
beer_brewery_merged_mean %>% slice(max_ibu_location_mean)
# This shows that the max ABV is in Astoria, Oregon


```

### Use the dropped for nan dataset
```{r}

# ABV
max(beer_brewery_merged_dropped_nan$ABV)
# This shows the max ABV is 0.125
max_abv_location_dropped <- which.max(beer_brewery_merged_dropped_nan$ABV)
# This shows that the max ABV is found in index 8
beer_brewery_merged_dropped_nan %>% slice(max_abv_location_dropped)
# This shows that the max ABV is in Louisville, KY

# IBU
max(beer_brewery_merged_dropped_nan$IBU)
# This shows the max ABV is 138
max_ibu_location_dropped <- which.max(beer_brewery_merged_dropped_nan$IBU)
# This shows that the max ABV is found in index 1134
beer_brewery_merged_dropped_nan %>% slice(max_ibu_location_dropped)
# This shows that the max ABV is in Astoria, Oregon


```


# Question 6
## Comment on the summary statistics and distribution of the ABV variable.
### Use the mean for nan dataset
```{r}

# Summary stats
summary(beer_brewery_merged_mean$ABV)

# We see the minimum is 0.001
# Max is 0.128
# Mean is 0.05977
# Median is 0.057
# First quarter is 0.05 
# Third quarter is 0.067

```
```{r}

# Histogram of ABV distribution
histogram(beer_brewery_merged_mean$ABV)

# Better histogram of ABV distribution
ABV_distribution_hist_mean <- ggplot(beer_brewery_merged_mean, aes(x = ABV)) +
  geom_histogram(fill = 'red', color = 'blue')

ABV_distribution_hist_mean
# We see that there is a decently normal distribution centered around 0.05
# There might be a tad bit of right skewness, but overall looks normal.


ABV_distribution_box_mean <- ggplot(beer_brewery_merged_mean, aes(x = ABV)) +
  geom_boxplot(fill = 'red', color = 'blue')

ABV_distribution_box_mean
# This box plot shows a few outliers on the right, which might 
# suggest a bit of a right skewness.

```

### Use the dropped for nan dataset
```{r}

# Summary stats
summary(beer_brewery_merged_dropped_nan$ABV)

# We see the minimum is 0.027
# Max is 0.125
# Mean is 0.05991
# Median is 0.057
# First quarter is 0.05 
# Third quarter is 0.068

```

```{r}

# Histogram of ABV distribution
histogram(beer_brewery_merged_dropped_nan$ABV)

# Better histogram of ABV distribution
ABV_distribution_hist_dropped <- ggplot(beer_brewery_merged_dropped_nan, aes(x = ABV)) +
  geom_histogram(fill = 'red', color = 'blue') +
  ggtitle('ABV Distribution')

ABV_distribution_hist_dropped
# We see that there is a decently normal distribution centered around 0.05
# There might be a tad bit of right skewness, but overall looks normal.


ABV_distribution_box_dropped <- ggplot(beer_brewery_merged_dropped_nan, aes(x = ABV)) +
  geom_boxplot(fill = 'red', color = 'blue') +
  ggtitle('ABV Distribution')

ABV_distribution_box_dropped
# This box plot shows a few outliers on the right, which might 
# suggest a bit of a right skewness.

```



# Question 7
## Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
### Use the mean for nan dataset
```{r}

# Draw a scatterplot to show the relationship between ABV and IBU
ggplot(beer_brewery_merged_mean, 
       aes(x = ABV, 
           y = IBU)) + 
  geom_point(shape = 18, 
             color = 'red') +
  geom_smooth(method = lm, 
              color = 'blue', 
              fill = 'blue')
# We see a moderate positive correlation between ABV and IBU. This shown 
# by the blue line increasing from left to right and bottom to top at a 
# moderate angle.

# Find the correlation coefficient 
cor(beer_brewery_merged_mean$ABV, beer_brewery_merged_mean$IBU)
# The correlation coefficient is 0.5200. This is an average positive correlation. 
# This suggests that ABV and IBU moderately influence the other. As one increases,
# the other also increases. 



```

### Use the dropped for nan dataset
```{r}

# Draw a scatterplot to show the relationship between ABV and IBU
ggplot(beer_brewery_merged_dropped_nan, 
       aes(x = ABV, 
           y = IBU)) + 
  geom_point(shape = 18, 
             color = 'red') +
  geom_smooth(method = lm, 
              color = 'blue', 
              fill = 'blue') +
  ggtitle('ABV vs. IBU')
# We see a moderate positive correlation between ABV and IBU. This shown 
# by the blue line increasing from left to right and bottom to top at a 
# moderate angle.

# Find the correlation coefficient 
cor(beer_brewery_merged_dropped_nan$ABV, beer_brewery_merged_dropped_nan$IBU)
# The correlation coefficient is 0.6706. This is an average positive correlation. 
# This suggests that ABV and IBU moderately influence the other. As one increases,
# the other also increases. 



```












































# Test Stuff

```{r}

# create a function that takes abbreviation and finds state name
abbr2state <- function(abbr){
  ab    <- tolower(c('AL', 'AK', 'AZ', 'KS', 'UT', 
                     'CO', 'CT', 'DE', 'FL', 'GA', 
                     'HI', 'ID', 'IL', 'IN', 'IA', 
                     'AR', 'KY', 'LA', 'ME', 'MD', 
                     'MA', 'MI', 'MN', 'MS', 'MO',
                     'MT', 'NE', 'NV', 'NH', 'NJ', 
                     'NM', 'NY', 'NC', 'ND', 'OH', 
                     'OK', 'OR', 'PA', 'RI', 'SC', 
                     'SD', 'TN', 'TX', 'CA', 'VT', 
                     'VA', 'WA', 'WV', 'WI', 'WY', 
                     'DC'))
  
  st    <- c('alabama', 'alaska', 'arizona', 'kansas',
             'utah', 'colorado', 'connecticut', 'delaware', 
             'florida', 'georgia', 'hawaii', 'idaho', 
             'illinois', 'indiana', 'iowa', 'arkansas',
             'kentucky', 'louisiana', 'maine',
             'maryland', 'massachusetts', 'michigan',
             'minnesota', 'mississippi', 'missouri',
             'montana', 'nebraska', 'nevada',
             'new Hampshire', 'new jersey', 'new mexico',
             'new York', 'north carolina', 'north dakota',
             'ohio', 'oklahoma', 'oregon', 'pennsylvania', 
             'rhode island', 'south carolina',
             'south Dakota', 'tennessee', 'texas',
             'california', 'vermont', 'virginia',
             'washington', 'west virginia', 'wisconsin',
             'wyoming', 'district of columbia')
  
  st[match(tolower(abbr), ab)]
}

```


```{r}

# Create a function that takes the abbreviation and returns the fips
abbr2fps <- function(abbr){
  ab    <- tolower(c('AL', 'AK', 'AZ', 'KS', 'UT', 
                     'CO', 'CT', 'DE', 'FL', 'GA',
                     'HI', 'ID', 'IL', 'IN', 'IA', 
                     'AR', 'KY', 'LA', 'ME', 'MD', 
                     'MA', 'MI', 'MN', 'MS', 'MO',
                     'MT', 'NE', 'NV', 'NH', 'NJ', 
                     'NM', 'NY', 'NC', 'ND', 'OH', 
                     'OK', 'OR', 'PA', 'RI', 'SC', 
                     'SD', 'TN', 'TX', 'CA', 'VT', 
                     'VA', 'WA', 'WV', 'WI', 'WY', 
                     'DC'))
  
  fps    <- c('01', '02', '04', '20',
              '49', '08', '09', '10', 
              '12', '13', '15', '16', 
              '17', '18', '19', '05',
              '21', '22', '23', '24', 
              '25', '26', '27', '28', 
              '29', '30', '31', '32',
              '33', '34', '35', '36', 
              '37', '38', '39', '40', 
              '41', '42', '44', '45',
              '46', '47', '48', '06', 
              '50', '51', '53', '54',
              '55', '56', '11')
  
  fps[match(tolower(abbr), ab)]
}

```


```{r}
# the abbreviation column had a leading space so we are deleting all extra spaces
ABV_Medians$State <- trimws(ABV_Medians$State, which = c('left'))
ABV_Medians$State
```


```{r}
# use the function to add column with fips
ABV_Medians$FPS<-abbr2fps(ABV_Medians$State)

ABV_Medians
```


```{r}
# Use the function to add column with full state name
ABV_Medians$Full_state<-abbr2state(ABV_Medians$State)

ABV_Medians
```

```{r}
# Try to plot on US map
 states <- map_data('state')

 sim_data_geo <- merge(states, ABV_Medians, by.x='region', by.y = 'Full_state')

 qplot(long, lat, data=sim_data_geo, geom='polygon', fill=Median, group=group)

 qplot(long, lat, data=sim_data_geo, geom='auto', fill=Median, group=group)


```






```{r}
ABV_Medians_dropped
ABV_Medians_dropped$State <- trimws(ABV_Medians_dropped$State, which = c('left'))
ABV_Medians_dropped$State
lookup = data.frame(abb = state.abb, State = state.name) #makes a data frame with State name and abbreviation. 
colnames(ABV_Medians_dropped)[1] = "abb" # Change Column Name
Acu2 = merge(ABV_Medians_dropped,lookup,"abb") # make one dataset with state names and abb
AcuMapData = count(Acu2,State) #count up the occurrence of each state. 
#AcuMapData = AcuMapData[-c(5,9,43),] #Shows contrast between other states better
colnames(AcuMapData)[2] = "Breweries" #change "n" to "Acuspikes"
AcuMapData$region <- tolower(AcuMapData$State)
AcuMapData2 = AcuMapData[-1]
states <- map_data("state")
map.df <- merge(states,AcuMapData2, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Breweries))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Acuspike Systems by State")+
coord_map()

```




# box plot for all states
## ABV
```{r}

# box plots for all states
beer_brewery_merged_mean %>%
  group_by(State) %>%
  mutate(mean_abv = mean(ABV, na.rm = T)) %>%
ggplot(aes(x = reorder(State, -mean_abv), y = ABV, fill = mean_abv)) + 
  geom_boxplot() + 
  labs(x = 'State', y = 'Beer ABV', title = 'ABV percentages by State') + 
  scale_y_continuous(breaks = seq(0, 12, 4)) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_fill_continuous(low = 'skyblue', high = 'blue', name = 'Mean ABV')



# box plots for all states
beer_brewery_merged_dropped_nan %>%
  group_by(State) %>%
  mutate(mean_abv = mean(ABV, na.rm = T)) %>%
ggplot(aes(x = reorder(State, -mean_abv), y = ABV, fill = mean_abv)) + 
  geom_boxplot() + 
  labs(x = 'State', y = 'Beer ABV', title = 'ABV percentages by State') + 
  scale_y_continuous(breaks = seq(0, 12, 4)) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_fill_continuous(low = 'skyblue', high = 'blue', name = 'Mean ABV')
```
## IBU
```{r}
# box plots for all states
beer_brewery_merged_mean %>%
  group_by(State) %>%
  mutate(mean_ibu = mean(IBU, na.rm = T)) %>%
ggplot(aes(x = reorder(State, -mean_ibu), y = ABV, fill = mean_ibu)) + 
  geom_boxplot() + 
  labs(x = 'State', y = 'Beer IBU', title = 'IBU percentages by State') + 
  scale_y_continuous(breaks = seq(0, 12, 4)) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_fill_continuous(low = 'skyblue', high = 'blue', name = 'Mean IBU')



# box plots for all states
beer_brewery_merged_dropped_nan %>%
  group_by(State) %>%
  mutate(mean_abv = mean(ABV, na.rm = T)) %>%
ggplot(aes(x = reorder(State, -mean_abv), y = ABV, fill = mean_abv)) + 
  geom_boxplot() + 
  labs(x = 'State', y = 'Beer ABV', title = 'ABV percentages by State') + 
  scale_y_continuous(breaks = seq(0, 12, 4)) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_fill_continuous(low = 'skyblue', high = 'blue', name = 'Mean ABV')


```


# jitter plot
```{r}

ggplot(beer_brewery_merged_dropped_nan, 
                aes(x = factor(Ounces), 
                    y = ABV,
                    col = IBU)) + 
  geom_jitter() + 
  labs(x = "Ounces", 
       y = "ABV", 
       title = "ABV by Ounces") +
  scale_color_gradient(low = "sky blue", 
                       high = "dark blue")



```


```{r}
ggplot(beer_brewery_merged_dropped_nan, 
                aes(x = factor(Ounces), 
                    y = IBU,
                    color = ABV)) + 
  geom_jitter() + 
  labs(x = "Ounces", 
       y = "IBU", 
       title = "IBU by Ounces") +
  scale_color_gradient(low = "sky blue", 
                       high = "dark blue")

```






# table for unique breweries and the mean abv ibu for each state

```{r}
install.packages("kableExtra")
library(kableExtra)


kable(beer_brewery_merged_dropped_nan %>% 
  group_by(State) %>% 
  summarise(n_unique = n_distinct(Name.y), 
  mean_abv = mean(ABV), mean_ibu = mean(IBU)) %>% 
  arrange(desc(n_unique)))



```

```{r}
ggplotly(beer_brewery_merged_dropped_nan %>% 
  group_by(State) %>% 
  summarise(n_unique = n_distinct(Name.y)) %>% 
  ggplot(aes(x = reorder(State, desc(n_unique)), y = n_unique)) + 
  geom_bar(stat="identity") + 
  theme_bw() + 
  labs(x = "State",y = "Unique Breweries", 
  title = "Number of Unique Breweries per State") +
    theme(axis.text.x = element_text(size=5, angle = 90)))
```




```{r}

# unique style of beers for each state
beer_brewery_merged_dropped_nan %>% 
  group_by(State) %>% 
  summarise(n_unique = n_distinct(Style)) %>% 
  arrange(desc(n_unique))

```



```{r}

ggplotly(beer_brewery_merged_dropped_nan %>% 
  group_by(State) %>% 
  summarise(n_unique = n_distinct(Style)) %>% 
  ggplot(aes(x = reorder(State, desc(n_unique)), y = n_unique)) + 
  geom_bar(stat="identity") + 
  theme_bw() + 
  labs(x = "State", y = "Unique Styles", 
  title = "Number of Unique Styles per State"))


```

```{r}
# table of mean stuff for each style

kable(beer_brewery_merged_dropped_nan %>% 
  group_by(Style) %>% 
  summarise(n = n(), mean_abv = mean(ABV), 
  mean_ibu = mean(IBU), mean_oz = mean(Ounces)) %>% 
  arrange(desc(n)))



```







